\documentclass{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{hyperref}
\usepackage{titling}
\pretitle{% add some rules
  \begin{flushleft}
    \huge\bfseries
}
\posttitle{%
  \end{flushleft}
}
\renewcommand\maketitlehooka{%
	\vspace{-10ex}
  \noindent\vrule height 2.5pt width \textwidth
}
\renewcommand\maketitlehookb{%
  \begin{flushleft}
   \Large\bfseries
    Cover Letter, 2021
  \end{flushleft}%
    \noindent\vrule height 2.5pt width \textwidth
}
\title{Shivesh Pathak}
\date{\vspace{-10ex}}
\begin{document}
\maketitle

\noindent Dear Professors of the John Hopkins University IQM,

I am writing to apply for the Condensed Matter Theory Postdoctoral Fellow position.
I am a graduate student at the University of Illinois at Urbana-Champaign and plan to graduate in August of 2021 working with my advisor Lucas K. Wagner.
My PhD thesis is entitled "Accurate low-energy states and interacting effective models using ab initio quantum Monte Carlo."

I am interested in this postdoctoral position because I have the skills required for this position and my research goals will contribute positively to the Institute's research goals.
As mentioned in my research statement and CV, I have a very strong background in computational and theoretical condensed matter physics, in particular for simulation and model development for strongly interacting quantum systems.
I have experience in building models and excited states for graphene-based systems, and in particular the strongly correlated twisted bilayer graphene system.
However, my research interests extend beyond these graphene systems as outlined in my future plans, and I can integrate well with the broader goal of the institute in furthering our understanding of strongly correlated electronic systems.
I also have experience in large, multi-research-group collaborations, with the QMC-HAMM collaboration, and will be able to extend my experience in effective communication with broad technical audiences to IQM.

I believe that this postdoctoral fellow position is an excellent opportunity for me to pursue my research goals, and I know that my research goals and expertise will help the institute accomplish their goals.
I look forward to speaking with all of you in the future, and I appreciate your consideration of my application.

\noindent Shivesh Pathak
\end{document}
